---
title: "lizard initial EDA"
author: "Kelly Fan"
date: '2023-10-18'
output: pdf_document
---
* clean duplicate rows
* two videos for each(almost) lizard; separate by condition: ob/noobs

# Setup
```{r message=FALSE}
#Load relevant libraries
library(dplyr)
library(plyr)
library(tidyr)
library(lme4)
# library(car)
library(lmerTest)
library(ggplot2)
library(emmeans)
library(tidyverse)
library(RRPP)
library(scales)
library(gridExtra)

#Clear memory
rm(list=ls(all = TRUE))
```

# Input data
```{r}
#Set working directory (assumes data file is available in same location as code)
current_path = rstudioapi::getActiveDocumentContext()$path 
setwd(dirname(current_path ))
#Uploading this ridiculous file and MorphData
DLCMasterData <- read.table("DLCMasterData.csv", header=T, sep = ",", as.is=T)
MorphData <- read.table("OWULizard_Morphology.csv", header=T, sep = ",", as.is=T)
OH_Podarcis_Data <- read.table("OH_Podarcis_Data - ExpLizards_2023.csv", header=T, sep = ",", as.is=T)
#Add new columns for gravidity status sex etc. to master doc
DLCMasterDataPlus <- merge(DLCMasterData, OH_Podarcis_Data, by = "Lizard_ID", all.x = T, all.y = T)
DLCMasterDataPlus$Sex <- as.factor(DLCMasterDataPlus$Sex)
#Videos were recorded at 120 fps; Re-scale x-axis accordingly
DLCMasterDataPlus$TimeSeconds <- DLCMasterDataPlus$bodyparts_coords/120
```

# Overview
```{r}
glimpse(DLCMasterDataPlus)
```

# Likelihood of m_ data
```{r}
detach(package:plyr)
DLCMasterDataPlus %>% 
  group_by(Condition) %>% 
  summarise(mean_mHead_llh = mean(m_Head_likelihood),
            mean_mSG_llh = mean(m_SG_likelihood),
            mean_mPG_llh = mean(m_PG_likelihood),
            mean_mTail_llh = mean(m_Tail_likelihood),
            mean_mURS_llh = mean(m_UR_S_likelihood),
            mean_mURZ_llh = mean(m_UR_Z_likelihood),
            mean_mURM_llh = mean(m_UR_M_likelihood),
            mean_mURT_llh = mean(m_UR_T_likelihood),
            mean_mLRS_llh = mean(m_LR_S_likelihood),
            mean_mLRZ_llh = mean(m_LR_Z_likelihood),
            mean_mLRM_llh = mean(m_LR_M_likelihood),
            mean_mLRT_llh = mean(m_LR_T_likelihood))
```
By this numerical summary, we observed that

* Trails that without obstacles have higher mean likelihoods
* m_Head (without obstacles) has the highest mean likelihood among all m_ series data, specifically around ~0.82, surpassing our cutoff of 0.75.
* Trails that with obstacles do not have mean likelihoods greater than 0.75

# Likelihood of NOT m_ data
```{r}
DLCMasterDataPlus %>% 
  group_by(Condition) %>% 
  summarise(mean_mHead_llh = mean(Head_likelihood),
            mean_SG_llh = mean(SG_likelihood),
            mean_PG_llh = mean(PG_likelihood),
            mean_Tail_llh = mean(Tail_likelihood),
            mean_URS_llh = mean(UR_S_likelihood),
            mean_URZ_llh = mean(UR_Z_likelihood),
            mean_URM_llh = mean(UR_M_likelihood),
            mean_URT_llh = mean(UR_T_likelihood),
            mean_LRS_llh = mean(LR_S_likelihood),
            mean_LRZ_llh = mean(LR_Z_likelihood),
            mean_LRM_llh = mean(LR_M_likelihood),
            mean_LRT_llh = mean(LR_T_likelihood))
```
There appears to be no discernible pattern in these non-m_ data, while their likelihood is consistently very low.

# Example: lizard Pm332
## Setup data frame
We created two separate data frames—one for the trail with obstacles and another for the trail without obstacles. Additionally, we removed entries with m_Head likelihood less than 0.75 from both data frames.
```{r}

pm332_ob <- DLCMasterDataPlus %>% 
  filter(Lizard_ID == 'Pm332' & Condition == 'Ob') %>% 
  filter(m_Head_likelihood > 0.75)
pm332_no <- DLCMasterDataPlus %>% 
  filter(Lizard_ID == 'Pm332' & Condition == 'NoObs') %>% 
  filter(m_Head_likelihood > 0.75)
```

## About m_head
### path
```{r}
# m_head path
mhead_ob <- ggplot(pm332_ob, aes(x=m_Head_x,y=m_Head_y, color=bodyparts_coords)) + geom_point()
mhead_no <- ggplot(pm332_no, aes(x=m_Head_x,y=m_Head_y, color=bodyparts_coords)) + geom_point()
grid.arrange(mhead_ob, mhead_no, ncol=2, nrow =1)
```
The path on the trail with obstacles appears to be more "curved" compared to the trail without obstacles. This observation makes sense, as they likely navigate around obstacles, resulting in a less straight and more curvier path.

### likelihood
```{r}
mean(pm332_ob$m_Head_likelihood) # with obstacles
mean(pm332_no$m_Head_likelihood) # no obstacles
```
The mean likelihood without obstacles is higher than likelihood with obstacles, consistent with our earlier observation in the numerical summary.

## Speed (in cm/s)
We computed the speed of lizard Pm332 by dividing the mirror length (provided as 42.18 cm) by the total time it took for lizard Pm332.
```{r}
pm332 <- DLCMasterDataPlus %>% 
  filter(Lizard_ID == 'Pm332')
mirror_length <- 42.18 #cm
mirror_length/pm332$TimeSeconds[nrow(pm332)] 
```

# Speed between Sex and Gravid status
## setup data frame
```{r}
mirror_length <- 42.18

# obtain each lizard's sex, gravid status and speed
lizard_ids <- levels(factor(DLCMasterDataPlus$Lizard_ID))
sexs <- c()
gravids <- c()
speeds <- c()
for (i in 1:length(lizard_ids)){
  id <- lizard_ids[i]
  lizard <- filter(DLCMasterDataPlus, Lizard_ID == id)
  sexs[i] <- lizard$Sex[1]
  gravids[i] <- lizard$Gravid[1]
  # obtain their speed
  n <- nrow(lizard)
  speed <- mirror_length/lizard$TimeSeconds[n]
  speeds[i] <- speed
}

# compile to a data frame
lizards <- data.frame(id = lizard_ids,
                      sex = as.factor(sexs), # 1: Female; 2: Male
                      gravid = gravids,
                      speed = speeds)
```

## box-plots comparing
```{r}
# female vs male
sex <- ggplot(lizards, aes(x=sex, y=speed))+
  geom_boxplot()+
  labs(title="Female(1) vs Male(2)")

# gravid vs nongravid
lizards_f <- lizards %>% 
  filter(!is.na(gravid))
gravid <- ggplot(lizards_f, aes(x=gravid, y=speed))+
  geom_boxplot()+
  labs(title="Nongravid(N) vs Gravid(Y)")

grid.arrange(sex, gravid, ncol=2, nrow =1)
```
By the box plots, we observed that

* The speed of all lizards are very similar (~5cm/s), regardless of their Sex and Gravid status. This aligns with our hypothesis: "males and females in different reproductive stages use their bodies differently while sprinting to achieve similar speeds despite their different proportions".
* The speed range of females is wider than that of males.
* The speed range of non-gravid lizards is wider than that of gravid ones.
* The speed of gravid lizard is slightly faster than non-gravid lizard.
